PInSoRo: A Dataset for Deep-learning of Human-Robot Social Interactions

### 2017-06-27 paper.tex
WIP

\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
    \hspace{1em}
    \includegraphics[width=0.4\linewidth]{child-child-env}
    \caption{The free play social interactions sandbox: two children interact in
    Da free-play situation, by drawing and manipulating items on a touchscreen.}
    \label{fig|setup}
\end{figure}

@@

We present a new experimental task based on \emph{free play} interactions. Pairs of
children (4-7 years old) are invited to freely draw and interact with items
displayed on an interactive table, telling and acting stories with no explicit
goal set in advance (Fig.~\ref{fig|setup}).

@@

The \emph{freeplay sandbox} follows the sandtray
paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
multitouch) is used as an interactive surface (\emph{sandtray}). The two children play together
by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
depicts a generic empty environment, with different symbolic zones (water,
grass, path, bushes...). The players do not have any particular task to
complete, they are simply invited to freely play.
is implemented wit To mitigate the technical challenges presented by free play interactions in
their most unconstrained version (like
high dimensionality and uncertainties), we have developped a \emph{free play
sandbox}, Fig.~\ref{fig|setup}:

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{sandbox}
    \caption{Example of a game situation: the two children `build'
    together a zoo. Items (animals, characters...) can be dragged over the whole
    play area, while the background picture can be painted over by picking a
    colour.}
    \label{fig|sandbox}
\end{figure}

children can engage in open-ended and
non-directive play situations, yet sufficiently
defined to be reproducible;

\subsubsection{Socio-cognitive Framing}

focus on abstract socio-cognitive facets (robot
perception and manipulation are simplified); well suited for qualitative and
quantitative analysis using metrics like Słowinski’s Individual Motor Signature
(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
interactions, with-me-ness (for assessment of co-engagment).

carefully framed: as far as possible, they ne-
cessitate only the cognitive capabilities that they ev-
idence (e.g. if they evidence a purely socio-linguistic
mechanism, they will not mandate complex action ca-
pabilities – they may benefit from it, though),

This point is especially important as it allows an
incremental path: a robot should be able to tackle one or
several of these task independently of the others, and a re-
search team may progressively extend their cognitive models
to incorporate more and more of the cognitive capabilities
required by the robot to address all of the tasks.

\subsubsection{Implementation}

The sandbox is implemented using mainly two frameworks: the Qt's \emph{QtQuick} framework
for the graphical interface of the game, and the \emph{Robot Operating System}
(ROS) for the modular implementation of the data processing and behaviour
generation pipelines.

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{rviz-sandtray}
    \caption{The freeplay sandbox, viewed at runtime within ROS RViz. The
    current background drawing is published on a regular image
    topic and computer vision is used to segment it into zones (visible in
    the central panel). The poses and bounding boxes of the interactive items
    are published as well, and turned into an occupancy map, used to plan the
    robot's arm motion.}
    \label{fig|rviz}
\end{figure}

@@

\begin{figure}
    \centering
    \includegraphics[width=0.9\linewidth]{3d-point-cloud-facial-features}
    \caption{Reconstructed 3D point cloud with 68 facial features high-lighted}
    \label{fig|pointcloud}
\end{figure}

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.4\linewidth]{setup_child_child_top}
    \hspace{1.5cm}
    \includegraphics[width=0.4\linewidth]{setup_child_robot_top}
    \caption{Experimental setup, child-child condition (left) and child-robot
    condition (right).}
\end{figure}

@@

\section{The PInSoRo Dataset}

\subsection{Structure}

As the data is recorded using ROS's bag files, it can be replayed in the exact
same conditions as it was during the recording. All the video streams use
calibrated cameras. While the raw RGB and depth streams are neither rectified
nor registered together (due to performance considerations during the
recording), it can be easily achieved as a post-process step, using standard ROS
tools (the ROS {\tt rgbd\_pipeline}). Pre-configured scripts are available
alongside the
dataset\footnote{Available online from
\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
republish registered streams and the corresponding 3D point-clouds (as seen in
Figure~\ref{fig|pointcloud}).

@@

\section{Discussion}

>>>>

\begin{figure}
    \centering
    \includegraphics[width=0.4\linewidth]{setup-child-child.png}
    \hspace{1em}
    \includegraphics[width=0.4\linewidth]{child-child-env}
    \caption{The free play social interactions sandbox: two children interact in
    a free-play situation, by drawing and manipulating items on a touchscreen.}
    \label{fig|freeplay}
\end{figure}

@@

We present a new experimental task based on \emph{free play} interactions. Pairs
of children (4-7 years old) are invited to freely draw and interact with items
displayed on an interactive table, telling and acting stories with no explicit
goal set in advance (Fig.~\ref{fig|freeplay}). The task is designed so that
children can engage in open-ended and non-directive play situations, yet
sufficiently well defined to be controlled, suitable for detailled recording,
and easily reproducible.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.8\linewidth]{sandbox}
    \caption{Example of a possible game situation. Items (animals,
    characters...) can be dragged over the whole play area, while the background
    picture can be painted over by picking a colour.}

    \label{fig|sandbox}
\end{figure}

The \emph{freeplay sandbox} follows the sandtray
paradigm~\cite{baxter2012touchscreen}: a large touchscreen (60cm $\times$ 33cm,
with multitouch support) is used as an interactive surface (\emph{sandtray}). The two children play together
by freely moving interactive items (Fig.~\ref{fig|sandbox}) on the surface. A background image
depicts a generic empty environment, with different symbolic colours (water,
grass, beach, bushes...). By drawing on top of the background picture, the
children can also alter this environment to their liking. The players do not have any particular task to
complete, they are simply invited to freely play for as long as they wish.

@@

%
%\paragraph{Socio-cognitive framing of the task}
%
%
%focus on abstract socio-cognitive facets (robot
%perception and manipulation are simplified); well suited for qualitative and
%quantitative analysis using metrics like Słowinski’s Individual Motor Signature
%(for behavioural alignment), Anderson's~\cite{anderson2004social} coding of children’s free-play
%interactions, with-me-ness (for assessment of co-engagment).
%
%carefully framed: as far as possible, they necessitate only the cognitive
%capabilities that they evidence (e.g. if they evidence a purely socio-linguistic
%mechanism, they will not mandate complex action capabilities – they may benefit
%from it, though),
%
%This point is especially important as it allows an incremental path: a robot
%should be able to tackle one or several of these task independently of the
%others, and a research team may progressively extend their cognitive models to
%incorporate more and more of the cognitive capabilities required by the robot to
%address all of the tasks.

\subsection{Dataset structure}

As the data is recorded using ROS's bag files, it can be replayed in the exact
same conditions as it was during the recording. All the video streams use
calibrated cameras. Only the raw RGB and depth video streams are recorded. While
the raw RGB and depth streams are neither rectified nor registered together (due
to performance considerations during the recording), it can be easily achieved
as a post-process step, using standard ROS tools (the ROS {\tt rgbd\_pipeline}).
Pre-configured scripts are available alongside the dataset\footnote{Available
online from
\url{https://github.com/severin-lemaignan/freeplay-sandbox-analysis}.} to
republish registered streams and the corresponding 3D point-clouds (as seen in
Fig.~\ref{fig|rviz}, left).

\subsection{Hardware and Software}

\paragraph{Hardware}
The sandtray is made of a 27" Samsung All-In-One computer running Ubuntu Linux
and equipped with a fast 1TB SSD hard-drive. The computer is held horizontal in
a custom aluminium frame standing 26cm above the ground
(Fig.~\ref{fig|freeplay}, right). All the cameras are plugged directly over
USB-3 to the computer. The computer performs all the data acquisition using ROS
Kinetic.

The children faces are recorded from two Intel RealSense SR300 RGB-D cameras
(short range, 0.2m to 1.2m, indoor RGB-D cameras), placed at the corners of the
the sandtray (Figs.~\ref{fig|rviz} and~\ref{fig|setup}), and tilted to face the
children. The cameras are rigidly mounted on custom 3D-printed brackets, and
their 6D pose with respect to the sandtray screen is precisely known and
recorded (this enables in particular accurate estimation of the children's head
pose).

Audio is recorded from the same SR300 cameras (one audio stream is recorded for
each child, from the camera facing her/him).

Finally, a third RGB camera (the RGB stream of a Microsoft Kinect One) records
the whole interaction setting. These video stream is intended to support human
coders while annontating the interaction, and is not precisely calibrated.

\begin{figure}[ht!]
    \centering
    \includegraphics[width=0.5\linewidth]{rviz-sandtray}
    \hspace{0.2em}
    \includegraphics[width=0.48\linewidth]{3d-point-cloud-facial-features}
    \caption{The freeplay sandbox, visualised at runtime with ROS RViz. The
    current background drawing is published on a regular image
    topic and computer vision is used to segment it into zones (visible in
    the central panel, right picture). The poses of the interactive items
    are published, as well as an estimate of the children's head pose.
    The left picture shows a reconstructed 3D point cloud of one child face,
    along with the detected 68 facial features.}
    \label{fig|rviz}
\end{figure}

\paragraph{Software}
The sandbox is implemented with two frameworks: the Qt's \emph{QtQuick} framework
for the graphical interface of the game, and the \emph{Robot Operating System}
(ROS) for the modular implementation of the data processing and data acquisition
pipelines. A dedicated bridge between QtQuick and ROS has been specifically
developped to enable the game interface to export the positions of every
iteractive items as they move, the background image whenever it is painted over,
and the children's touches. The game interface is open-source and available online:
\url{https://github.com/severin-lemaignan/freeplay-sandbox-qt}.

By relying on ROS for the data acquisition, real-time monitoring of the
interaction is also possible (Fig.~\ref{fig|rviz}).
The ROS data acquisition pipeline is open-source as well, and available from
\url{https://github.com/severin-lemaignan/freeplay-sandbox-ros}.

Finally, we have developped a web-based \emph{supervisor} that permits the
experimenter to remotely start/stop the ROS nodes and the game GUI, as well as to
record annotations during the experiment. The supervisor ensures that the exact
same recording procedure (detailled in the next section) is followed for every
participants. The supervisor is available online as well:
\url{https://github.com/severin-lemaignan/freeplay-sandbox-supervisor}.

@@

@@

\begin{figure}[htbp]
    \centering
    \includegraphics[width=0.4\linewidth]{setup_child_child_top}
    \hspace{1.5cm}
    \includegraphics[width=0.4\linewidth]{setup_child_robot_top}
    \caption{Experimental setup, child-child condition (left) and child-robot
    condition (right).}
    \label{fig|setup}
\end{figure}

@@

@@

\section{Applications}
